Supplemental Material: Multidimensional analysis and detection of informative features in human brain white matter

### 2021-04-28 reports/manuscript/paper_changes.tex
Move figure captions to the end.

\begin{figure}[b!]
    % \vspace{0.25cm}
    % \rlap{
    %     \hspace{-0.25em}
    %     \includegraphics[width=0.98\columnwidth]{methods-brains.pdf}
    % }
    % \vspace{-8.4cm}
    % \includegraphics[width=\columnwidth]{method-quad.pdf}
    {\phantomsubcaption\label{fig:methods:tractogram}}
    {\phantomsubcaption\label{fig:methods:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile:fa}}
    {\phantomsubcaption\label{fig:methods:group-structure}}
    \caption{\textbf{Tractometry data flow}
        \label{fig:methods}
        \textbf{(a)} Whole brain tractography generates streamlines approximating
        the trajectories of white matter connections.
        \textbf{(b)} Tractometry classifies these streamlines into anatomical bundles.
        In this case, we show the left corticospinal tract (CSTL)
        \added{and the left arcuate fasciculus (ARCL)}
        over a mid-saggital anatomical slice.
        \deleted{\textbf{(c)}} \replaced{Tract profiling}{Tractometry} further extracts bundle profiles,
        quantifications of various diffusion metrics along the length of the
        fiber bundle. Here, we show one subject's fractional anisotropy (FA)
        profile for \added{\textbf{(c)}} the CSTL \added{and \textbf{(d)} the ARCL}.
        \textbf{(d)} the phenotypical target data and \replaced{tract profile}{tractometric}
        features can be organized into a linear model, $\hat{y} = \mathbf{X}
        \hat{\beta}$. The feature matrix $\mathbf{X}$ is color-coded
        to reveal a natural group structure: the left (orange) group
        contains $k$ features from the CSTL, the middle (green) group
        contains $k$ features from the left cingulum cingulate (CGCL),
        and the right (blue) group
        contains $k$ features from the \replaced{ARCL}{left arcuate (ARCL)}.
        The coefficients in $\hat{\beta}$ follow the same natural grouping.
    }
\end{figure}

@@

\begin{figure}[t!]
    % \vspace{3.0cm}
    % \rlap{
    %     \includegraphics[width=0.33\textwidth]{als_gpca_coefs.pdf}
    % }
    % \vspace{-7.0cm}
    % \includegraphics[width=\textwidth]{als_gpca_results.pdf}
    {\phantomsubcaption\label{fig:class-results:probs}}
    {\phantomsubcaption\label{fig:class-results:coefs3d}}
    {\phantomsubcaption\label{fig:class-results:tract-profiles}}
    {\phantomsubcaption\label{fig:class-results:false-positives}}
    {\phantomsubcaption\label{fig:class-results:false-negatives}}
    \caption{%
        {\bf \added{PCR-}SGL accurately and interpretably predicts ALS diagnosis.}
        \label{fig:class-results}
        {\bf (a)} Classification probabilities for ALS diagnosis, with
        controls on the left, patients on the right, predicted controls in
        blue, and predicted patients in orange. That is, orange dots on the left
        represent false positives, while blue dots on the right represent
        false negatives. We achieve {\protect\alsAccuracy}\% accuracy with an
        ROC AUC of {\protect\alsRocAuc}.
        {\bf (b)} \added{PCR-}SGL coefficients are presented on the core fibers of major
        fiber bundles. They exhibit high group sparsity and are concentrated
        in the FA of the \replaced{corticospinal tract (CST)}{CST}.
        The brain is oriented with the right hemisphere
        in the foreground and anterior to the right of the page. The CSTL, CSTR,
        \replaced{callosum forceps anterior (CFA), left arcuate (ARCL), and right arcuate (ARCR)}{CFA}
        bundles are indicated for orientation.
        {\bf (c)} \added{PCR-}SGL identifies three portions of the CST as important,
        where $\hat{\beta}$ (dashed line, right axis) has large values. These
        are centered around nodes 30, 65, and 90, corresponding to locations
        of substantial differences in FA between the ALS and control groups
        (shaded areas indicates standard error of the mean).
        {\bf (d)} Bundle profiles for false positive classifications. Line
        colors correspond to the marker edge color in the top left plot.
        These individuals have reduced FA in the CST portions which SGL
        identified as important. Their misclassification is coherent with the
        feature importance and the group differences in FA.
        {\bf (e)} Individual bundle profiles for false negative
        classifications. These individuals have bundle profiles which
        oscillate between the group means.
    }
\end{figure}

@@

\begin{figure}[t!]
    % \vspace{3.0cm}
    % \rlap{
    %     \hspace{0.775cm}
    %     \includegraphics[width=0.26\textwidth]{wh_coefs.pdf}
    %     \hspace{0.5em}
    %     \includegraphics[width=0.26\textwidth]{hbn_coefs.pdf}
    %     \hspace{0.5em}
    %     \includegraphics[width=0.26\textwidth]{camcan_coefs.pdf}
    % }
    % \vspace{-6.1cm}
    % \includegraphics[width=\textwidth]{regression_scatter.pdf}
    {\phantomsubcaption\label{fig:age-results:wh-scatter}}
    {\phantomsubcaption\label{fig:age-results:hbn-scatter}}
    {\phantomsubcaption\label{fig:age-results:cc-scatter}}
    {\phantomsubcaption\label{fig:age-results:wh-coefs3d}}
    {\phantomsubcaption\label{fig:age-results:hbn-coefs3d}}
    {\phantomsubcaption\label{fig:age-results:cc-coefs3d}}
    {\phantomsubcaption\label{fig:age-results:wh-profile}}
    {\phantomsubcaption\label{fig:age-results:hbn-profile}}
    {\phantomsubcaption\label{fig:age-results:cc-profile}}
    \caption{%
        {\bf Predicting age with tractometry and SGL.}
        \label{fig:age-results}
        {\bf (top)} The predicted age vs. true age of each individual from the test
        splits (i.e., when each subject's data was held out in fitting the
        model) for the {\bf (a)} WH, {\bf (b)} HBN, and {\bf (c)} Cam-CAN
        datasets; an accurate prediction falls close to the $y=x$ line
        (dashed). The mean absolute error (MAE) and coefficient of
        determination $R^2$ are presented in the lower right of each scatter
        plot.
        {\bf (middle)} Feature importance for predicting age from
        \replaced{tract profile}{tractometry} in
        the {\bf (d)} WH, {\bf (e)} HBN, and {\bf (f)} Cam-CAN datasets.
        The orientation of the
        brain is that same as in \cref{fig:class-results:coefs3d}, however because
        the coefficients exhibit high global sparsity (as opposed to group
        sparsity), we plot the mean of the absolute value of $\hat{\beta}$
        for each bundle on the core fiber. The global distrubution of the
        $\hat{\beta}$ coefficients reflects the fact that aging is not
        confined to a single white matter bundle.
        {\bf (bottom)} Age quintile bundle profiles for the {\bf (g)} WH,
        {\bf (h)} HBN, and {\bf (i)} Cam-CAN datasets.
    }
\end{figure}

@@

\begin{figure}[t!]
    % \includegraphics[width=\textwidth]{model_performance.pdf}
    \caption{%
        {\bf \added{Model performance across all datasets.}}
        \label{fig:model-performance}
        \added{Each panel shows model performance measured on the test set for each}
        \added{cross-validation split, with each black dot representing a split, box}
        \added{plots representing the quartiles, and white diamonds representing the mean}
        \added{performance. The $y$-scale varies in each subplot.}
        \added{\textbf{(a)} Accuracy of test set predictions for the}
        \added{ALS dataset. Because group differences}
        \added{in ALS diagnosis are mostly confined to a single bundle, the group}
        \added{structure-preserving methods, SGL and PCR-SGL, outperform the other}
        \added{models.}
        \added{The remaining frames show coefficient of determination, \protect$R^2$ in test sets}
        \added{for the \textbf{(b)} WH, \textbf{(c)} HBN, and \textbf{(c)} Cam-CAN}
        \added{datasets. Because aging affects the white matter globally, group structure-blind}
        \added{methods like elastic net and PCR Lasso perform well. Nonetheless, the SGL}
        \added{models show competitve predictive performance, adapting to a problem where}
        \added{group structure is not as informative. PCR-SGL performs poorly in this regime}
        \added{because its initial group-wise PC projection destroys between bundle covariance.}
        \added{The bundle-mean lasso performs poorly, demonstrating the value of along-tract profiling.}
    }
\end{figure}
\added{Model performance across all four datasets and all six model types is summarized}
\added{in \cref{fig:model-performance}. In contrast to the ALS classification case, the PCR Lasso performs}
\added{competitively in age regression, suggesting that aging is a significant source of global variance}
\added{in the tract profiles. On the other hand, the PCR-SGL models destroy cross-bundle covariance information}
\added{in their initial group-wise PC projection step (see \cref{eq:svd}), limiting performance}
\added{for globally distributed phenomena. Conversely, the SGL models are able to adapt to this regime.}

@@

\bibliography{paper}
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for
% step-by-step instructions.
\end{document}

>>>>

@@

@@

@@

\added{Model performance across all four datasets and all six model types is summarized}
\added{in \cref{fig:model-performance}. In contrast to the ALS classification case, the PCR Lasso performs}
\added{competitively in age regression, suggesting that aging is a significant source of global variance}
\added{in the tract profiles. On the other hand, the PCR-SGL models destroy cross-bundle covariance information}
\added{in their initial group-wise PC projection step (see \cref{eq:svd}), limiting performance}
\added{for globally distributed phenomena. Conversely, the SGL models are able to adapt to this regime.}

@@

\bibliography{paper}
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for
% step-by-step instructions.

\newpage

\section*{Figure legends}

\begin{figure}[b!]
    % \vspace{0.25cm}
    % \rlap{
    %     \hspace{-0.25em}
    %     \includegraphics[width=0.98\columnwidth]{methods-brains.pdf}
    % }
    % \vspace{-8.4cm}
    % \includegraphics[width=\columnwidth]{method-quad.pdf}
    {\phantomsubcaption\label{fig:methods:tractogram}}
    {\phantomsubcaption\label{fig:methods:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile:cst}}
    {\phantomsubcaption\label{fig:methods:tract-profile:fa}}
    {\phantomsubcaption\label{fig:methods:group-structure}}
    \phantomcaption{\label{fig:methods}}
\end{figure}

\begin{figure}[b!]
    % \vspace{3.0cm}
    % \rlap{
    %     \includegraphics[width=0.33\textwidth]{als_gpca_coefs.pdf}
    % }
    % \vspace{-7.0cm}
    % \includegraphics[width=\textwidth]{als_gpca_results.pdf}
    {\phantomsubcaption\label{fig:class-results:probs}}
    {\phantomsubcaption\label{fig:class-results:coefs3d}}
    {\phantomsubcaption\label{fig:class-results:tract-profiles}}
    {\phantomsubcaption\label{fig:class-results:false-positives}}
    {\phantomsubcaption\label{fig:class-results:false-negatives}}
    \phantomcaption{%
    \label{fig:class-results}
    }
\end{figure}

\begin{figure}[b!]
    % \vspace{3.0cm}
    % \rlap{
    %     \hspace{0.775cm}
    %     \includegraphics[width=0.26\textwidth]{wh_coefs.pdf}
    %     \hspace{0.5em}
    %     \includegraphics[width=0.26\textwidth]{hbn_coefs.pdf}
    %     \hspace{0.5em}
    %     \includegraphics[width=0.26\textwidth]{camcan_coefs.pdf}
    % }
    % \vspace{-6.1cm}
    % \includegraphics[width=\textwidth]{regression_scatter.pdf}
    {\phantomsubcaption\label{fig:age-results:wh-scatter}}
    {\phantomsubcaption\label{fig:age-results:hbn-scatter}}
    {\phantomsubcaption\label{fig:age-results:cc-scatter}}
    {\phantomsubcaption\label{fig:age-results:wh-coefs3d}}
    {\phantomsubcaption\label{fig:age-results:hbn-coefs3d}}
    {\phantomsubcaption\label{fig:age-results:cc-coefs3d}}
    {\phantomsubcaption\label{fig:age-results:wh-profile}}
    {\phantomsubcaption\label{fig:age-results:hbn-profile}}
    {\phantomsubcaption\label{fig:age-results:cc-profile}}
    \phantomcaption{%
    \label{fig:age-results}
    }
\end{figure}

\begin{figure}[b!]
    % \includegraphics[width=\textwidth]{model_performance.pdf}
    \phantomcaption{%
    \label{fig:model-performance}
    }
\end{figure}

\begin{figure}[h!]
    % \includegraphics[width=0.75\textwidth]{nested-cross-validation.pdf}
    \phantomcaption{
    \label{fig:nested-cross-val}
    }
\end{figure}

\paragraph{Figure 1: Tractometry data flow}
\textbf{(a)} Whole brain tractography generates streamlines approximating
    the trajectories of white matter connections.
    \textbf{(b)} Tractometry classifies these streamlines into anatomical bundles.
    In this case, we show the left corticospinal tract (CSTL)
    \added{and the left arcuate fasciculus (ARCL)}
    over a mid-saggital anatomical slice.
    \deleted{\textbf{(c)}} \replaced{Tract profiling}{Tractometry} further extracts bundle profiles,
    quantifications of various diffusion metrics along the length of the
    fiber bundle. Here, we show one subject's fractional anisotropy (FA)
    profile for \added{\textbf{(c)}} the CSTL \added{and \textbf{(d)} the ARCL}.
    \textbf{(d)} the phenotypical target data and \replaced{tract profile}{tractometric}
    features can be organized into a linear model, $\hat{y} = \mathbf{X}
    \hat{\beta}$. The feature matrix $\mathbf{X}$ is color-coded
    to reveal a natural group structure: the left (orange) group
    contains $k$ features from the CSTL, the middle (green) group
    contains $k$ features from the left cingulum cingulate (CGCL),
    and the right (blue) group
    contains $k$ features from the \replaced{ARCL}{left arcuate (ARCL)}.
    The coefficients in $\hat{\beta}$ follow the same natural grouping. \added{Panels (a) and (b) are adapted from \url{https://figshare.com/articles/figure/example_tractography-segmentation/14485350}, and reproduced under the CC-BY license (\url{https://creativecommons.org/licenses/by/4.0/})}

\paragraph{Figure 2: \added{PCR-}SGL accurately and interpretably predicts ALS diagnosis.}
{\bf (a)} Classification probabilities for ALS diagnosis, with
controls on the left, patients on the right, predicted controls in
blue, and predicted patients in orange. That is, orange dots on the left
represent false positives, while blue dots on the right represent
false negatives. We achieve {\protect\alsAccuracy}\% accuracy with an
ROC AUC of {\protect\alsRocAuc}.
{\bf (b)} \added{PCR-}SGL coefficients are presented on the core fibers of major
fiber bundles. They exhibit high group sparsity and are concentrated
in the FA of the \replaced{corticospinal tract (CST)}{CST}.
The brain is oriented with the right hemisphere
in the foreground and anterior to the right of the page. The CSTL, CSTR,
\replaced{callosum forceps anterior (CFA), left arcuate (ARCL), and right arcuate (ARCR)}{CFA}
bundles are indicated for orientation.
{\bf (c)} \added{PCR-}SGL identifies three portions of the CST as important,
where $\hat{\beta}$ (dashed line, right axis) has large values. These
are centered around nodes 30, 65, and 90, corresponding to locations
of substantial differences in FA between the ALS and control groups
(shaded areas indicates standard error of the mean).
{\bf (d)} Bundle profiles for false positive classifications. Line
colors correspond to the marker edge color in the top left plot.
These individuals have reduced FA in the CST portions which SGL
identified as important. Their misclassification is coherent with the
feature importance and the group differences in FA.
{\bf (e)} Individual bundle profiles for false negative
classifications. These individuals have bundle profiles which
oscillate between the group means.

\paragraph{Figure 3: Predicting age with tractometry and SGL.}
{\bf (top)} The predicted age vs. true age of each individual from the test
splits (i.e., when each subject's data was held out in fitting the
model) for the {\bf (a)} WH, {\bf (b)} HBN, and {\bf (c)} Cam-CAN
datasets; an accurate prediction falls close to the $y=x$ line
(dashed). The mean absolute error (MAE) and coefficient of
determination $R^2$ are presented in the lower right of each scatter
plot.
{\bf (middle)} Feature importance for predicting age from
\replaced{tract profile}{tractometry} in
the {\bf (d)} WH, {\bf (e)} HBN, and {\bf (f)} Cam-CAN datasets.
The orientation of the
brain is that same as in \cref{fig:class-results:coefs3d}, however because
the coefficients exhibit high global sparsity (as opposed to group
sparsity), we plot the mean of the absolute value of $\hat{\beta}$
for each bundle on the core fiber. The global distrubution of the
$\hat{\beta}$ coefficients reflects the fact that aging is not
confined to a single white matter bundle.
{\bf (bottom)} Age quintile bundle profiles for the {\bf (g)} WH,
{\bf (h)} HBN, and {\bf (i)} Cam-CAN datasets.

\paragraph{\added{Figure 4: Model performance across all datasets.}}
\added{Each panel shows model performance measured on the test set for each}
\added{cross-validation split, with each black dot representing a split, box}
\added{plots representing the quartiles, and white diamonds representing the mean}
\added{performance. The $y$-scale varies in each subplot.}
\added{\textbf{(a)} Accuracy of test set predictions for the}
\added{ALS dataset. Because group differences}
\added{in ALS diagnosis are mostly confined to a single bundle, the group}
\added{structure-preserving methods, SGL and PCR-SGL, outperform the other}
\added{models.}
\added{The remaining frames show coefficient of determination, \protect$R^2$ in test sets}
\added{for the \textbf{(b)} WH, \textbf{(c)} HBN, and \textbf{(c)} Cam-CAN}
\added{datasets. Because aging affects the white matter globally, group structure-blind}
\added{methods like elastic net and PCR Lasso perform well. Nonetheless, the SGL}
\added{models show competitve predictive performance, adapting to a problem where}
\added{group structure is not as informative. PCR-SGL performs poorly in this regime}
\added{because its initial group-wise PC projection destroys between bundle covariance.}
\added{The bundle-mean lasso performs poorly, demonstrating the value of along-tract profiling.}

\paragraph{Figure 5: Nested cross-validation.}
We evaluate model quality using a nested $k$-fold cross validation
scheme. At level-0, the input data is decomposed into $k_0$ shuffled
groups and optimal hyperparameters are found for the level-0 training
set. To avoid overfitting, the optimal hyperparameters are themselves
evaluated using a cross-validation scheme taking place at level-1 of
the decomposition, where each level-0 training set is further
decomposed into $k_1 = 3$ shuffled groups. In the classification
case, the training and test splits are stratified by diagnosis. For
the ALS and WH data, $k_0 = 10$, while for the HBN and Cam-CAN data,
$k_0 = 5$.

\end{document}
